\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-11-10 icra-2021.tex
updated the text

@@

Many scenarios involving unmanned aerial vehicles (UAVs) such as precision agriculture, search and rescue, and surveillance require high autonomy but have limited energy budgets. A typical example of a mission in these scenarios is a UAV flying a trajectory and performing some on-board computations with little human interaction. For instance, the UAV might detect ground patterns and inform other ground-based actors. Energy-aware dynamic mission planning finds optimal tradeoffs between the trajectory, computations, and energy requirements. Current generic planning solutions for UAVs flying outdoors do not plan the mission dynamically, nor are energy-aware. They are often semi-autonomous; the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy.

@@

The proposed algorithm unfolds as follows. The user specifies an initial mission plan--the trajectory and the tasks--along with an energy budget and some parameters. The Quality of Service (QoS) computations parameters are relative to the tasks, and trajectory-explicit equations (TEEs) adjustments parameters to the path. The algorithm replans the initial mission plan as the energy budget changes due to atmospheric interferences. In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE. Our goal is to complete the mission with the highest possible computations and adjustments as the UAV flies and its batteries drain.

The algorithm optimizes computations built as software components that run on a UAV's companion embedded device. Then, it optimizes adjustments--a way to alter the trajectory--and guides the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. Uncertainty accounts for atmospheric interference. E.g., a fixed-wing UAV drifts due to windy weather. Periodicity accounts for the mission's repetitive patterns. Indeed UAV scenarios often iterate over a set of tasks and trajectories (e.g., a monitoring or search and rescue mission). The algorithm addresses the latter by modeling the energy with Fourier analysis. The mission is periodic, so we expect the energy to approximately evolve periodically. It addresses uncertainty with a state estimator and outputs the control input (adjustments and computations) using output model predictive control (MPC).

In the spirit of reducing costs and resources, we showcase the algorithm using the problem of dynamic mission planning for a precision agriculture fixed-wing UAV. Such a scenario is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The mission is structured as follows. Trajectory-wise, the UAV flies in circles and lines covering a polygon. Computationally-wise, it detects obstacles using a convolutional neural network (CNN), and informs grounded mobile robots employed for harvesting. The algorithm plans the mission controlling the processing rate and the radius of the circles (affecting the distance between the lines). Figure~\ref{fig:overview} shows an initial slice of such a mission. The UAV first heads to a circular TEE with a given radius, which is later reduced as, e.g., windy weather requires the adjustment of the control input to avoid battery depletion. Computations significantly impact the battery, with a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations (see Section~\ref{sec:experimental}).

@@

\begin{defn}[Stage and mission]\label{def:mission}
  A \emph{stage} $\mathcal{M}_i$ at time instant $k$ of a mission $\mathcal{M}$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \mathcal{M}_i:=\{\langle\varphi_i(\mathbf{p}_k,c&_{i,1},\dots,c_{i,\rho}),\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\rangle\\\mid
      \exists\,\,\mathbf{p}_k,\,&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\psi_j(s_{i,j})\in\mathbb{S}_{i,j}\,\forall j\in[\sigma]^+\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEEs set, and $\mathbb{S}_{i,j}:=[\underline{s}_{i,j},\overline{s}_{i,j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th task QoS set. $\mathbf{p}_k:=(x,y)$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$. The parameters of the TEE $\varphi_i$ are the point and the adjustments (Subsection~\ref{sec:model}). The parameters of the tasks $\psi_1,\dots,\psi_\sigma$ are the computations (Subsection~\ref{sec:computations-model}).

>>>>

  @inproceedings{seewald2020beyond,
    title={Beyond Traditional Energy Planning: the Weight of Computations in Planetary Exploration},
    author={Seewald, Adam},
    booktitle={IROS Workshop on Planetary Exploration Robots: Challenges and Opportunities (PLANROBO20)},
    organization={ETH Zurich, Department of Mechanical and Process Engineering},
    url={https://adamseew.bitbucket.io/short/beyond2020}
  @inproceedings{lindemann2005smoothly,
    title={Smoothly blending vector fields for global robot navigation},
    author={Lindemann, Stephen R and LaValle, Steven M},
    booktitle={Proceedings of the 44th IEEE Conference on Decision and Control},
    pages={3553--3559},
  @article{gonccalves2010vector,
    title={Vector fields for robot navigation along time-varying curves in $n$-dimensions},
    author={Gon{\c{c}}alves, Vin{\'\i}cius M and Pimenta, Luciano CA and Maia, Carlos A and Dutra, Bruno CO and Pereira, Guilherme AS},
    journal={IEEE Transactions on Robotics},
    volume={26},
    pages={647--659},
  @inproceedings{panagou2014motion,
    title={Motion planning and collision avoidance using navigation vector fields},
    author={Panagou, Dimitra},
    booktitle={2014 IEEE International Conference on Robotics and Automation (ICRA)},
    pages={2513--2518},
  @inproceedings{zhou2014vector,
    title={Vector field following for quadrotors using differential flatness},
    author={Zhou, Dingjiang and Schwager, Mac},
    booktitle={2014 IEEE International Conference on Robotics and Automation (ICRA)},
    pages={6567--6572},
  @article{kapitanyuk2017guiding,
    title={A guiding vector-field algorithm for path-following control of nonholonomic mobile robots},
    author={Kapitanyuk, Yuri A and Proskurnikov, Anton V and Cao, Ming},
    journal={IEEE Transactions on Control Systems Technology},
    volume={26},
    pages={1372--1385},

@@

Many scenarios involving unmanned aerial vehicles (UAVs) such as precision agriculture, search and rescue, and surveillance require high autonomy but have limited energy budgets. A typical example of a mission in these scenarios is a UAV flying a trajectory and performing some on-board computations with little human interaction. For instance, the UAV might detect ground patterns and inform other ground-based actors. Energy-aware dynamic mission planning finds optimal tradeoffs between the trajectory, computations, and energy requirements. However, current generic planning solutions for UAVs flying outdoors do not plan the mission dynamically, nor are energy-aware. They are often semi-autonomous; the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV mission planning problem.

@@

The proposed algorithm contributes to the UAV mission planning problem as follows. The user specifies an initial mission plan--the trajectory and the tasks--along with an energy budget and some parameters: the Quality of Service (QoS) computations are relative to the tasks. The trajectory-explicit equations (TEEs) adjustments are relative to the path. The computations specify a given task's computational level. The adjustments are a way to alter the TEE. The algorithm replans the initial mission plan as the energy budget changes due to atmospheric interferences. The goal is to complete the mission with the highest possible computations and adjustments as the UAV flies and its batteries drain. 
The algorithm plans both tasks and trajectory unlike most of the past planning algorithms literature. It guides the UAV using a vector field~\cite{de2017guidance} that converges smoothly to the planned trajectory. The vector field for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding}. 
In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE.

The algorithm contributes to the planning literature further with a periodic energy model that accounts for the uncertainty (atmospheric interference). Mission periodicity is usually present due to the mission's repetitive patterns~\cite{seewald2020mechanical}. Indeed UAV scenarios often iterate over a set of tasks and trajectories (e.g., monitoring or search and rescue). The mission is periodic, so we expect the energy to approximately evolve periodically. The observation applies broadly; repetitive patterns have been observed in scenarios as far as space rovers~\cite{seewald2020beyond}. The algorithm addresses the periodicity through Fourier analysis. The uncertainty with a state estimator. It outputs the control input (adjustments and computations) using output model predictive control (MPC).

In the spirit of reducing costs and resources, we showcase the algorithm using the problem of dynamic mission planning for a precision agriculture fixed-wing UAV. Precision agriculture is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The mission is structured as follows. Trajectory-wise, the UAV flies in circles and lines covering a polygon. Computationally-wise, it detects obstacles using a convolutional neural network (CNN) and informs grounded mobile robots employed for harvesting. The algorithm plans the mission; it controls the processing rate and the radius of the circles (affecting the distance between the lines). Figure~\ref{fig:overview} shows an initial slice of such a mission. The UAV first heads to a circular TEE with a given radius, which is later reduced as, e.g., windy weather requires the adjustment of the control input to avoid battery depletion. Computations significantly impact the battery, with a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations (see Section~\ref{sec:experimental}).

@@

\begin{defn}[Stage and mission]\label{def:mission}
  A \emph{stage} $\mathcal{M}_i$ at time instant $k$ of a mission $\mathcal{M}$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \mathcal{M}_i:=\{\langle\varphi_i(\mathbf{p}_k,c&_{i,1},\dots,c_{i,\rho}),\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\rangle\\\mid
      \exists\,\,\mathbf{p}_k,\,&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\psi_j(s_{i,j})\in\mathbb{S}_{i,j}\,\forall j\in[\sigma]^+\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEEs set, and $\mathbb{S}_{i,j}:=[\underline{s}_{i,j},\overline{s}_{i,j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th task QoS set. $\mathbf{p}_k:=(x,y)$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$. The parameters of the TEE $\varphi_i$ are the point and the adjustments (Subsection~\ref{sec:model}). The parameters of the tasks $\psi_1,\dots,\psi_\sigma$ are the computations (Subsection~\ref{sec:computations-model}). Note that while the TEE can differ for each stage, the tasks remain the same. However, the user can inhibit or enable a task varying QoS set.
